---
title: "Lab 6"
author: "Katie Munster"
date: "11/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(palmerpenguins)
library(broom)
library(equatiomatic)
```

## Rank based test exxample

Mann-Whitney U (non-parametric alternative to an unpaired 2-sample t-test).

H0: Our samples are drawn from populations with equal ranks (or medians)
HA: Ranks are not equal (or, medians are not equal)

### Create samples

`set.seed()` to create "pseudorandom, but reproducible samples"

```{r}
set.seed(1414) # setting the seed gives us all the same starting point. we all get the same output
gp_1 <- sample.int(20, size = 15, replace = TRUE)
# 20 is the highest possible value
# with sample replacement so it is a random sample

set.seed(1424)
gp_2 <- sample.int(30, size = 15, replace = TRUE)
# run gp_2 in the console to see its values. we all got the same random sample
```

Take a quick look at the distributions using the `hist()` function. 

```{r}
hist(gp_1)
# not screaming normality. only 15 observations but doesnt seem like it is definitely from a normal distribution
hist(gp_2)
# also does not scream normal distribution. looks more like an uniform distribution

# leads me to think it might be worth to do a more conservative rank-based test for these values
```

Try a t-test so we can compare the results:

```{r}
ttest <- t.test(gp_1, gp_2)
# unpaired t test
# If these data were paired, I can just add the argument 'paired = "TRUE"' within the t.test() function.
# t test is  a question about means differences

# This p-value means there is a 19.8% chance that there are two samples with means at least this different if they come from populations of the same mean.
```

This p-value means that IF our samples are drawn from populations with the same mean (i.e. if the null hypothesis is true), there is `r round(ttest$p.value, 2)` probability of finding sample means *at least as different* as mine by random chance.

[A more comprehensive explanation of difference between groups...actually means, data spread, effect size]...there is no significant difference (t(df) = t-statistic value, p = ).

(degrees of freedom is a measure of pool sample size)

### Mann Whitney U rank based test

We use the `wilcox.test()` function to perform a two-sample rank-based test.

```{r}
mwu <- wilcox.test(gp_1, gp_2)
# type mwu into console to see the values
# W or U is the statistic value
#?wilcox.test in the console to learn more about the variables
```

This p-value means that IF our samples are drawn from populations with the same median (i.e. if the null hypothesis is true), there is `r round(ttest$p.value, 2)` probability of finding sample medians *at least as different* as mine by random chance.

No significant difference in medians (Mann-Whitney U; U = `r mwu$statistic`, p = `r round(mwu$p.value, 2)`)

The medians of group A (M = 22.4 g) and group B (M = 31.2 g) did not differ significantly (Mann Whitney U, ...)


